Fail Faster

### 2025-03-15 paper/paper.tex
intro

\jwc{``Property based testing is a race against time...''}

>>>>

\jwc{``Property based testing is a race against time...''}
\\
\tr{Property-based testing (PBT) uses generators of random data to produce inputs to a
system under test, which is then checked against a set of properties the system is
 expected to satisfy.  In practice, PBT is used as a lightweight, ad hoc approach 
 to software correctness, aimed at detecting bugs within a short testing 
 window---typically 50 milliseconds to 30 seconds. Given this practical constraint, 
 it is 
 imperative that inputs be generated as fast as possible; in other words, 
 generators must be treated as performance-sensitive code.}

\tr{The locus for improvement in this area is \textit{generator libraries}: if programmers are 
 provided with the tools to write fast generators, then they do not themselves 
 need to be performance experts. However, existing libraries fail to achieve good performance 
 due to two major sources of inefficiency. The first is \textit{random sampling}, or the process
 of selecting random values. The first is random sampling, or the process of selecting random values. 
 The second is abstraction: while generator libraries are designed for flexibility and expressiveness, 
 this comes at the cost of performance, resulting in frequent boxing and unboxing and unnecessary 
 allocations (which lead to costly garbage collection pauses). In strict functional languages, 
 constructs like monadic bind introduce function call overhead, requiring each step in a generator 
 sequence to allocate closures and thread state through function applications.
 }
